### 2023-05-01 chapter0/chapter0.tex
Clarified wording around Zarin2019 summary in introduction

IDRs, however, challenge this sequence-dependent model of protein structure and function. Because they do not make stable contacts between residues which establish a fixed structure, IDRs are not generally constrained to maintain a specific sequence of residues. Thus, while there are exceptions, many IDRs evolve extremely rapidly, and related IDRs are therefore not easily identified by their sequence. There is growing evidence, though, that IDRs evolve under a different set of constraints. Because the composition and patterning of residues in an IDR dictates its conformational class, many distinct sequences can yield similar conformational ensembles. Furthermore, because modification and binding sites in IDRs are usually fewer than ten residues, their interaction interfaces are compact and can occur in multiple positions without compromising function~\cite{Tompa2014}. Thus, rather than conserving specific sequences, IDRs are hypothesized to conserve distributed ``molecular features'' associated with those sequences. By the mid 2010s several studies had demonstrated evidence of such constraint in the flexibility~\cite{Daughdrill2007}, chemical composition~\cite{Moesa2012}, net charge~\cite{Zarin2017}, or charge distribution of IDRs~\cite{Beh2012}. While these initial studies were generally restricted to specific features or proteins, in 2019 Zarin \textit{et al.} showed various features from a comprehensive set of IDR-associated properties were conserved in IDRs across the entire yeast proteome~\cite{Zarin2019}. By comparing the observed values of these features against those generated under a simulated model of evolution, they clustered IDRs by their ``evolutionary signatures,'' \textit{i.e.} patterns of conserved features. Furthermore, they identified specific biological functions associated with these groups, which for the first time provided a global view of the relationship between sequence and function in IDRs.

>>>>

IDRs, however, challenge this sequence-dependent model of protein structure and function. Because they do not make stable contacts between residues which establish a fixed structure, IDRs are not generally constrained to maintain a specific sequence of residues. Thus, while there are exceptions, many IDRs evolve extremely rapidly, and related IDRs are therefore not easily identified by their sequence. There is growing evidence, though, that IDRs evolve under a different set of constraints. Because the composition and patterning of residues in an IDR dictates its conformational class, many distinct sequences can yield similar conformational ensembles. Furthermore, because modification and binding sites in IDRs are usually fewer than ten residues, their interaction interfaces are compact and can occur in multiple positions without compromising function~\cite{Tompa2014}. Thus, rather than conserving specific sequences, IDRs are hypothesized to conserve distributed ``molecular features'' associated with those sequences. By the mid 2010s several studies had demonstrated evidence of such constraint in the flexibility~\cite{Daughdrill2007}, chemical composition~\cite{Moesa2012}, net charge~\cite{Zarin2017}, or charge distribution of IDRs~\cite{Beh2012}. These initial studies demonstrating evidence of constraint were generally restricted to specific features or proteins. However, by comparing the observed values of various IDR-associated properties against those generated under a simulated model of evolution, in 2019 Zarin \textit{et al.} showed most IDRs across the entire yeast proteome contain conserved features. Furthermore, they identified clusters of IDRs with common ``evolutionary signatures,'' \textit{i.e.} patterns of conserved features, which were associated with specific biological functions. This analysis for the first time provided a global view of the relationship between sequence and function in IDRs.

### 2023-05-01 chapter2/chapter2.tex
Clarified wording around Zarin2019 summary in introduction

While the studies which had initially demonstrated evidence of constraint were generally restricted to specific features or proteins, in 2019 Zarin \textit{et al.} showed various features from a comprehensive set of IDR-associated properties were conserved in IDRs across the entire yeast proteome~\cite{Zarin2019}. By comparing the observed values of these features against those generated under a simulated model of evolution, they clustered IDRs by their ``evolutionary signatures,'' \textit{i.e.} patterns of conserved features. Furthermore, they identified specific biological functions associated with these groups, which for the first time provided a global view of the relationship between sequence and function in IDRs. A follow-up analysis expanded on this initial finding by applying techniques from machine learning and statistics to predict the functions of individual IDRs using their evolutionary signatures~\cite{Zarin2021}.

>>>>

The initial studies demonstrating evidence of constraint were generally restricted to specific features or proteins. However, by comparing the observed values of various IDR-associated properties against those generated under a simulated model of evolution, Zarin \textit{et al.} showed most IDRs across the entire yeast proteome contain conserved features. Furthermore, they identified clusters of IDRs with common ``evolutionary signatures,'' \textit{i.e.} patterns of conserved features, which were associated with specific biological functions. This analysis for the first time provided a global view of the relationship between sequence and function in IDRs. A follow-up study then expanded on this initial finding by applying techniques from machine learning and statistics to predict the functions of individual IDRs using their evolutionary signatures~\cite{Zarin2021}.
